### 2020-03-04 doc/icfp20/music.tex

[TODO] Maybe need to explain degrees, instead of duration and notes.

@@

A \emph{pitch} tells us how high or low a sound is.
We represent pitches as natural numbers, but throughout the paper,
we use a more readable notation \texttt{name octave}, where
\texttt{name} ranges over \texttt{c}, \texttt{d}, \texttt{e}, etc.,
and \texttt{octave} denotes which octave the pitch belongs to.
For instance, the middle C is represented as \texttt{c 5}.

@@

@@

\emph{Duration} denotes an unspecified unit of time during which
a sound or silence lasts.
We represent duration as a natural number, and use this value to
calculate the absolute length when the music is played at a specific
tempo.
As an example, the duration constant \texttt{whole} has value
\texttt{16} and corresponds to 4 seconds if the tempo is 60 beats
per minute.

@@

\paragraph{Notes}

Combining pitches and duration gives us \emph{notes}.
We represent notes as a datatype \texttt{Note} with two constructors:
\texttt{tone} for notes with sound and \texttt{rest} for those without.
For example, \texttt{tone whole (c 5)} is a whole note whose sound
is the middle C.

@@

An \emph{interval} represents the difference in pitch between
two notes.
There are 13 kinds of interval within an octave
(Figure~\ref{fig:interval}), and these intervals can be classified from
several different perspectives:
(i) major or minor; (ii) consonant or dissonant; and (iii) perfect
or imperfect.
We represent intervals as natural numbers, but in this paper, we
refer to them using more informative names such as \texttt{maj3}
and \texttt{per8}.

@@

A \emph{chord} refers to a set of simultaneously sounding notes.  In
this paper we are particularly interested in \emph{triads}, which are
chords consisting of three notes separated by the 3rd intervals (i.e.,
\texttt{maj3} and \texttt{min3}).  We define triads as a datatype
\texttt{Triad} with roman numeral constructors representing the scale
degree of the lowest note (called the \emph{root}). The other two
notes are then members of the scale a third and a fifth above the
root. For instance, \texttt{I} in C major represents the triad C-E-G.

>>>>

@@

The most basic concept in melody is \emph{pitches}, which tell us
how high or low a tone is.
We represent pitches as natural numbers, but throughout the paper,
we use a more readable notation \texttt{name octave}, where
\texttt{name} ranges over \texttt{c}, \texttt{d}, \texttt{e}, etc.,
and \texttt{octave} denotes which octave the pitch belongs to.
For instance, the middle C is represented as \texttt{c 5}.

@@

\paragraph{Scale Degrees}

While pitches uniquely determine the position of a tone on e.g.
a piano keyboard, \emph{scale degrees} stand for the relative position
of a tone in a scale.
For instance, a diatonic scale consists of seven scale degrees, which
corresponds to C, D, E, F, G, A, B in the case of C major.
We represent scale degrees as a datatype with seven constructors
\texttt{d1} - \texttt{d7}.

@@

For rhythm, the fundamental concept is \emph{duration}, an
unspecified unit of time during which a sound or silence lasts.
We represent duration as a natural number, and use this value to
calculate the absolute length when the music is played at a specific
tempo.
As an example, the duration constant \texttt{whole} has value
\texttt{16} and corresponds to 4 seconds if the tempo is 60 beats
per minute.

@@

% \paragraph{Notes}

% Combining pitches and duration gives us \emph{notes}.
% We represent notes as a datatype \texttt{Note} with two constructors:
% \texttt{tone} for notes with sound and \texttt{rest} for those without.
% For example, \texttt{tone whole (c 5)} is a whole note whose sound
% is the middle C.

@@

An \emph{interval} represents the difference in pitch between
two notes.
There are 13 kinds of interval within an octave
(Figure~\ref{fig:interval}), and these intervals can be classified from
several different perspectives:
(i) major or minor; (ii) consonant or dissonant; and (iii) perfect
or imperfect.
We represent intervals as natural numbers, but in this paper, we
refer to them using names such as \texttt{maj3} and \texttt{per8},
which are more reader-friendly.

@@

A \emph{chord} refers to a set of simultaneously sounding notes.
In this paper we are particularly interested in \emph{triads}, which are
chords consisting of three notes separated by the 3rd intervals (i.e.,
\texttt{maj3} and \texttt{min3}).
We define triads as a datatype with roman numeral constructors
representing the scale degree of the lowest note (called the \emph{root}).
The other two notes are then members of the scale a third and a fifth
above the root.
For instance, the triad \texttt{I} in C major corresponds to the chord
C-E-G.
Triads can also be arranged in a way that a non-root note appears as
the lowest note.
Such arrangement is called \emph{inversion}; if a triad is not applied
inversion, it is in \emph{root position}.
